The four postulates of quantum mechanics are three

### 2020-12-01 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
After review with Christine

	\begin{defn}[Composite systems]\label{def_comp}
		Let A and B be two systems. The composite system C of A and B is formed by the simple collection of those and only those two systems, in the sense that it satisfies the following two requirements.
		\begin{enumerate}
			\item Every preparation of both subsystems is a preparation of the
			composite. Formally, let $\pj{\mathcal{C}}$ be the state space
			for C, there exists a map (not yet specified)
			$M:\pj{\mathcal{A}}\times\pj{\mathcal{B}}\to\pj{\mathcal{C}}$
			such that, for any compatible pair of (pure) states $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$, the proposition $\pj{a} \wedge \pj{b}$ is equivalent to the (pure) state $M(\pj{a},\pj{b}) \in \pj{\mathcal{C}}$ where $M$ returns the state of the composite system where the subsystems were prepared in the given states. In other words, $\pj{a} \wedge \pj{b}$ and $M(\pj{a},\pj{b})$ correspond to the same event in probability space\footnote{We will end up proving that the map $M$ leads to the tensor product.}.
			\item For every preparation of the composite, local projective measurements must have at least one
			outcome with non-zero probability. Formally, for every $\pj{c} \in \pj{\mathcal{C}}$, we can find at least $\pj{a} \in \pj{\mathcal{A}}$ and $\pj{b} \in \pj{\mathcal{B}}$ such that $P(\pj{a} \wedge \pj{b}|\pj{c})\neq 0$. 
		\end{enumerate}
		It is important to understand that these requirements are necessary. Requirement 1 ensures that the composite system is well defined at least when the components are prepared independently. Conceptually, this ensures that the composite system contains all the properties of the components. Note that superselection rules or other restrictions may prevent the independent preparation of all possible pairs (e.g.~two fermions cannot be jointly prepared in the same state). The tensor product is recovered only when all pairs are compatible. Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components. Violation of the second requirement would mean that some composite states would not define all the properties of the subsystems. That is, while the systems $A$ and $B$ by itself would define an observable $O_A$ and $O_B$, when put grouped together all values would be assigned zero probability; those observables no longer exist. In this case, the nature of the system would have changed so radically we would no longer call it a composite system.
	\end{defn}

@@

		In simple terms, for a Hilbert space, the conjugate
		vector space is equivalent to the dual space, so we
		could equivalently choose one or the other. For
		example, Schr\"odinger in the papers in which he
		introduces the Schr\"odinger equation, he writes it
		with both signs, as the choice of sign of the imaginary part of the wave function is arbitrary: one sign refers to
		the Hilbert space, the other to the dual space, namely
		to the complex-conjugate wave function \cite{sch}. So we can think of the anti-linear map as one that preserves the inner product but maps ket vectors into bra vectors. Looking ahead, the above result does not exclude a composition map similar to the tensor product, but that maps the kets of one or both subsystems into bras in the composite system. This would only make the representation of the composite physical system more complicated, as we need to keep track of the different conventions in the different subspaces. Therefore, without changing the physics, we can always mathematically redefine the second space so that the resulting map is linear. With this in mind, we will assume that the map between the spaces is linear, which will in turn lead to identifying the tensor product as a unique composition map.

@@

	The above proof relies only on the independent preparation of subsystem and not their measurement. However, during the review process, an anonymous referee contributed a sketch for a proof that shows very directly that the state and measurement postulates also imply independence of measurements. The insight is that the mixture created by all possible measurement outcomes on $B$ must behave overall as a pure state on $A$. Since pure states are extreme points, this can only happen if every measurement outcome on $B$ leaves $A$ in a pure state, which makes the probability factorize.

	\begin{prop}
		Let $\pj{\psi}$ and $\pj{\phi}$ be two preparations for $A$ and $B$. Let $\pj{a}$ and $\pj{b}$ two measurements on the respective systems. Then $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi}) P(\pj{b} | \pj{\phi})$.
	\end{prop}
	\begin{proof}
		Consider $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi})$. We can imagine first performing the measurement on $B$ and then conditioning the result on $A$. We have $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{b}, \pj{\psi} \wedge \pj{\phi}) P(\pj{b} | \pj{\psi} \wedge \pj{\phi})$ where by ``$\pj{b}, \pj{\psi} \wedge \pj{\phi}$'' we mean that the systems were prepared in $\pj{\psi}$ and $\pj{\phi}$ respectively and $\pj{b}$ was measured. Any preparation on $A$ can be expressed with a mixed state, and any measurement on $B$ depends only on $\phi$, so we have $P(\pj{a} | \pj{b}, \pj{\psi} \wedge \pj{\phi}) P(\pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \rho(\pj{b}, \pj{\psi}, \pj{\phi})) P(\pj{b} | \pj{\phi})$. We also must have $\sum_b P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi})$.

>>>>

	\begin{defn}[Composite systems]\label{def_comp}
		Let A and B be two systems. The composite system C of A and B is formed by the simple collection of those and only those two systems, in the sense that it satisfies the following two requirements.
		\begin{enumerate}
			\item Every preparation of both subsystems is a preparation of the
			composite. Formally, let $\pj{\mathcal{C}}$ be the state space
			for C, there exists a map (not yet specified)
			$M:\pj{\mathcal{A}}\times\pj{\mathcal{B}}\to\pj{\mathcal{C}}$
			such that, for any compatible pair of (pure) states $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$, the proposition $\pj{a} \wedge \pj{b}$ is equivalent to the (pure) state $M(\pj{a},\pj{b}) \in \pj{\mathcal{C}}$ where $M$ returns the state of the composite system where the subsystems were prepared in the given states. In other words, $\pj{a} \wedge \pj{b}$ and $M(\pj{a},\pj{b})$ correspond to the same event in probability space\footnote{We will end up proving that the map $M$ leads to the tensor product.}.
			\item For every preparation of the composite, local projective measurements must have at least one
			outcome with non-zero probability. Formally, for every $\pj{c} \in \pj{\mathcal{C}}$, we can find at least $\pj{a} \in \pj{\mathcal{A}}$ and $\pj{b} \in \pj{\mathcal{B}}$ such that $P(\pj{a} \wedge \pj{b}|\pj{c})\neq 0$. 
		\end{enumerate}
		It is important to understand that these requirements are necessary. Requirement 1 ensures that the composite system is well defined at least when the components are prepared independently. Conceptually, this ensures that the composite system contains all the properties of the components. Note that superselection rules or other restrictions may prevent the independent preparation of all possible pairs (e.g.~two fermions cannot be jointly prepared in the same state). The tensor product is recovered only when all pairs are compatible. Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components. Violation of the second requirement would mean that some composite states would not define all the properties of the subsystems. That is, while the systems $A$ and $B$ by themselves would define observables $O_A$ and $O_B$, when grouped together all values would be assigned zero probability (XXX); those observables no longer exist. In this case, the nature of the system would have changed so radically we would no longer call it a composite system.
	\end{defn}

@@

		In simple terms, for a Hilbert space, the conjugate
		vector space is equivalent to the dual space, so we
		could equivalently choose one or the other. For
		example, Schr\"odinger, (XXX) in the papers in which he
		introduces the Schr\"odinger equation, (XXX) writes it
		with both signs, as the choice of sign of the imaginary part of the wave function is arbitrary: one sign refers to
		the Hilbert space, the other to the dual space, namely
		to the complex-conjugate wave function \cite{sch}. So we can think of the anti-linear map as one that preserves the inner product but maps ket vectors into bra vectors. Looking ahead, the above result does not exclude a composition map similar to the tensor product, but that maps the kets of one or both subsystems into bras in the composite system. This would only make the representation of the composite physical system more complicated, as we need to keep track of the different conventions in the different subspaces. Therefore, without changing the physics, we can always mathematically redefine the second space so that the resulting map is linear. With this in mind, we will assume that the map between the spaces is linear, which will in turn lead to identifying the tensor product as a unique composition map.

@@

	The above proof relies only on the independent preparation of subsystems and not their measurements. (XXX) However, during the review process, an anonymous referee contributed a sketch for a proof that shows very directly that the state and measurement postulates also imply independence of measurements. The insight is that the mixture created by all possible measurement outcomes on $B$ must behave overall as a pure state on $A$. Since pure states are extreme points, this can only happen if every measurement outcome on $B$ leaves $A$ in a pure state, which makes the probability factorize.

	\begin{prop}
		Let $\pj{\psi}$ and $\pj{\phi}$ be two preparations for $A$ and $B$. Let $\pj{a}$ and $\pj{b}$ be (XXX) two measurements on the respective systems. Then $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi}) P(\pj{b} | \pj{\phi})$.
	\end{prop}
	\begin{proof}
		Consider $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi})$. We can imagine first performing the measurement on $B$ and then conditioning the result on $A$. We have $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{b}, \pj{\psi} \wedge \pj{\phi}) P(\pj{b} | \pj{\psi} \wedge \pj{\phi})$ where by ``$\pj{b}, \pj{\psi} \wedge \pj{\phi}$'' we mean that the systems were prepared in $\pj{\psi}$ and $\pj{\phi}$ respectively and $\pj{b}$ was measured. Any preparation on $A$ can be expressed with a mixed state, and any measurement on $B$ depends only on $\phi$, so we have $P(\pj{a} | \pj{b}, \pj{\psi} \wedge \pj{\phi}) P(\pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \rho(\pj{b}, \pj{\psi}, \pj{\phi})) P(\pj{b} | \pj{\phi})$. We also must have $\sum_b P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi})$.

### 2020-12-01 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
After review with Christine

	We start from the natural definition of a composite system as the set
	of two (or more) quantum systems. The composite system is therefore
	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing
		else}. The first key insight is that the first two postulates of
	quantum theory (introduced below) already assume that the preparation
	of one system is independent from the preparation of another
	(statistical independence). In fact, we cannot even talk about a
	system in the first place if we cannot characterize it independently.
	The second key insight is that, since all measurement outcomes and all
	states of the subsystems plus the composite live in the same
	probability space, we will automatically have a map $M$ that for a
	pair of subsystems' states gives us the composite state that
	corresponds to the statistically independent case. These insights are
	enough to characterize mathematically the state space of the
	composite: the linearity given by the Hilbert space, together with the
	fact that the composite system is fully described by the observables
	of $A$ and $B$, allows to extend the construction to the composite
	states that do not satisfy statistical independence (the entangled
	states). So the work consists of two interrelated efforts: a physical
	argument that starts from the first two postulates and leads to the
	necessary existence of the composition map $M$ and its properties
	together with a formal argument that shows how $M$ leads to the tensor
	product.

@@

	This map $M$ acts on the state spaces of the subsystems. Each pure
	state is identified by a ray $\pj{\psi}$, a subspace of the system's
	Hilbert space spanned by all vectors $\psi$ differing by their
	(nonzero) modulus and phase: a two-dimensional subspace (a plane). In
	the same way, constraining the observable $X$ to a particular outcome
	value $x_0$ means identifying the subspace spanned by all
	non-normalized eigenvectors $|x_0\>$ of arbitrary phase such that
	$X | x_0 \> = x_0 |x_0\>$. As we need to map states and observables
	between subsystems and composite system, the map $M$ derived from the
	probability space is really a map between subspaces, namely the
	projective spaces. We then use projective geometry to show that the
	map $M$ on the projective spaces corresponds to a map $m$ on the
	vector spaces which is the tensor product.

	This is done by showing that preparation independence and statistical
	independence of the systems imply three conditions on the map $m$:
	(H1)~totality: the map is defined on all states of the subsystems;
	(H2)~bilinearity: the map is bilinear thanks to the fundamental
	theorem of projective geometry; (H3)~span surjectivity: the span of
	the image of map coincides with the full composite Hilbert space.  We
	then prove that, if these three conditions H1, H2 and H3 hold, then
	the map $m$ is the tensor product, namely the Hilbert space of the
	composite system is a tensor product of the components: the tensor
	product ``postulate'', which hence loses its status of a postulate. An
	overview of all these logical implications is given in
	Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
	argument, including all the physical arguments outlined above. The
	supplementary material contains the mathematical details.

@@

	We start with the axiomatization of quantum mechanics based on the
	following postulates
	(e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The pure state
	of a system is described by a ray $\pj{\psi}$ corresponding to a set
	of non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
	system's observable properties are described by self-adjoint operators
	acting on that space; (b)~The probability that a measurement of a
	property $X$, described by the operator with spectral decomposition
	$\sum_{x,i}x|x_i\>\<x_i|/|\<x_i|x_i\>|$ ($i$ a degeneracy index),
	returns a value $x$ given that the system is in state $\pj{\psi}$ is
	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2/|\<x_i|x_i\>\<\psi|\psi\>|$ (Born
	rule). (c)~The state space of a composite system is given by the
	tensor product of the spaces of the component systems; (d)~The time
	evolution of an isolated system is described by a unitary operator
	acting on a vector representing the system state,
	$|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or, equivalently, by the
	Schr\"odinger equation. The rest of quantum theory can be derived from
	these axioms. While some axiomatizations introduce further postulates,
	we will be using only (a) and (b) to derive (c), so the above are
	sufficient to our aims.

@@

	Step 2: use the universal property. The tensor product is uniquely
	characterized, up to isomorphism, by a universal property regarding
	bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
	tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
	map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the
	property than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$
	factors through $T$ uniquely.  This means that there exists a {\em
		unique} $\hat i$, dependent on $m$, such that $\hat i \circ T=m$.
	In other words, the following diagram commutes:
	\begin{center}
		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{i}"] \\
			& \mathcal{C}
		\end{tikzcd}
	\end{center}
	Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a
	bilinear operator (property H2), thanks to the universal property of
	the tensor product we can find a unique linear operator
	$\hat{i} : \mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that
	$m(a, b) = \hat i(a \otimes b)$. The set $\{ \hat i(a_i\otimes b_j)$
	with $|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and
	${\cal B}\}$ forms a basis for $\cal C$, since
	$\hat i(a_i\otimes b_j)=m(a_i,b_j)$ and we have shown above that the
	latter is a basis.  Thus,
	\begin{align} 
	&\<\hat i(a_i\otimes b_j)|\hat i(a_k\otimes b_\ell)
	\>_{\mathcal{C}}=\<m(a_i,
	b_j)|m(a_k,b_\ell)\>_\mathcal{C} \nonumber\\& =
	\delta_{ik}\delta_{j\ell}
	= \<a_i\otimes
	b_j| a_k \otimes b_\ell\>_{\otimes},
	\labell{ecco}\; 
	\end{align}
	where we used the orthonormality of the bases and the fact that
	$|a_i\otimes b_j\>$ is a basis of the tensor product space
	${\cal A}\otimes{\cal B}$. Since the function $\hat{i}$ is a linear
	function that maps an orthonormal basis of ${\cal A}\otimes{\cal B}$
	to an orthonormal basis of $\cal C$, $\hat{i}$ is a an isomorphism (a
	bijection that preserves the mathematical structure) between
	${\cal A}\otimes{\cal B}$ and $\cal C$. As
	${\cal C}\cong{\cal A}\otimes{\cal B}$ are isomorphic as Hilbert
	spaces, they are mathematically equivalent: we can loosely say that
	$\hat i$ is the identity. So we can directly use the
	tensor product to represent the composite state space. This means that
	the map $m : \cal A \times \cal B \to \cal C$ is equivalent to the map
	$\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$ in the
	sense that $m \circ \hat i^{-1} = \otimes$.$\square$

@@

	\vskip 1\baselineskip A few comments on the proof: it is based on the
	universal property of the tensor product, which uniquely characterizes
	it. In step 1 we show that the bilinear map $m$ maps subsystems' bases
	into the composite system basis. We also know that there exists a
	tensor product map $ {T}=\otimes$ that can compose the vectors in
	$\cal A$ and $\cal B$.  In step 2 we use the universal property: since
	$m$ is a bilinear map, we are assured that there exists a unique
	$\hat i$ such that $\hat i \circ T=m$. Since we show that $\hat i$ is
	an isomorphism, then $\hat{i}$ bijectively maps vectors in $\cal C$
	onto vectors in the tensor product space. Namely $m={T}=\otimes$.

>>>>

Let us first provide a conceptual overview of our approach. (XXX) We start from the natural definition of a composite system as the set
of two (or more) quantum systems. The composite system is therefore
made of system $A$ {\em and} (joined with) system $B$ and {\em nothing
  else}. The first key insight is that the first two postulates of
quantum theory (introduced below) already assume that the preparation
of one system is independent from the preparation of another
(statistical independence). In fact, we cannot even talk about a
system in the first place if we cannot characterize it independently.
The second key insight is that, since all measurement outcomes and all
states of the subsystems plus the composite live in the same
probability space, we will automatically have a map $M$ that for a
pair of subsystems' states gives us the composite state that
corresponds to the statistically independent case. These insights are
enough to characterize mathematically the state space of the
composite: the linearity given by the Hilbert space, together with the
fact that the composite system is fully described by the observables
of $A$ and $B$, allows us (XXX) to extend the construction to the composite
states that do not satisfy statistical independence (the entangled
states). So the work consists of two interrelated efforts: a physical
argument that starts from the first two postulates and leads to the
necessary existence of the composition map $M$ and its properties
together with a formal argument that shows how $M$ leads to the tensor
product.

@@

This map $M$ acts on the state spaces of the subsystems. Each pure
state is identified by a ray $\pj{\psi}$, a subspace of the system's
Hilbert space comprising (XXX) all vectors $\psi$ differing by their
(nonzero) modulus and phase: a one-dimensional complex subspace (a complex plane). In
the same way, constraining the observable $X$ to a particular outcome
value $x_0$ means identifying the subspace comprising (XXX) all
non-normalized eigenvectors $|x_0\>$ of arbitrary phase such that
$X | x_0 \> = x_0 |x_0\>$. As we need to map states and observables
between subsystems and composite system, the map $M$ derived from the
probability space is really a map between subspaces, namely the
projective spaces. We then use projective geometry to show that the
map $M$ on the projective spaces corresponds to a map $m$ on the
vector spaces which is the tensor product.

This is done by showing that preparation independence and statistical
independence of the systems imply three conditions on the map $m$:
(H1)~totality: the map is defined on all states of the subsystems;
(H2)~bilinearity: the map is bilinear thanks to the fundamental
theorem of projective geometry; (H3)~span surjectivity: the span of
the image of the map coincides with the full composite Hilbert space.  We
then prove that, if these three conditions H1, H2 and H3 hold, then
the map $m$ is the tensor product, namely the Hilbert space of the
composite system is the (XXX) tensor product of the components' Hilbert spaces (XXX). The tensor
product ``postulate'' hence loses its status of a postulate. (XXX) An
overview of all these logical implications is given in
Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
argument, including all the physical arguments outlined above. The
supplementary material contains the mathematical details.

@@

We start with the axiomatization of quantum mechanics based on the
following postulates
(e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The pure state
of a system is described by a ray $\pj{\psi}$ corresponding to a set
of non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
system's observable properties are described by self-adjoint operators
acting on that space; (b)~The probability that a measurement of a
property $X$, described by the operator with spectral decomposition
$\sum_{x,i}x\frac{|x_i\>\<x_i|}{\<x_i|x_i\>}$ ($i$ a degeneracy index),
returns a value $x$ given that the system is in state $\pj{\psi}$ is
$p(x|\psi)=\sum_i\frac{|\<\psi|x_i\>|^2}{\<x_i|x_i\>\<\psi|\psi\>}$ (Born
rule). (c)~The state space of a composite system is given by the
tensor product of the spaces of the component systems; (d)~The time
evolution of an isolated system is described by a unitary operator
acting on a vector representing the system state,
$|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or, equivalently, by the
Schr\"odinger equation. The rest of quantum theory can be derived from
these axioms. While some axiomatizations introduce further postulates,
we will be using only (a) and (b) to derive (c), so the above are
sufficient to our aims.

@@

Step 2: use the universal property. The tensor product is uniquely
characterized, up to isomorphism, by a universal property regarding
bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the
property than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$
factors through $T$ uniquely.  This means that there exists a {\em
  unique} $I$, dependent on $m$, such that $I \circ T=m$.
In other words, the following diagram commutes:
\begin{center}
		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "I"] \\
			& \mathcal{C}
		\end{tikzcd}
\end{center}
Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a
bilinear operator (property H2), thanks to the universal property of
the tensor product we can find a unique linear operator
$I : \mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that
$m(a, b) = I(a \otimes b)$. The set $\{ I(a_i\otimes b_j)$
with $|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and
${\cal B}\}$ forms a basis for $\cal C$, since
$I(a_i\otimes b_j)=m(a_i,b_j)$ and we have shown above that the
latter is a basis.  Thus,
\begin{align} 
  &\<I(a_i\otimes b_j)|I(a_k\otimes b_\ell)
    \>_{\mathcal{C}}=\<m(a_i,
    b_j)|m(a_k,b_\ell)\>_\mathcal{C} \nonumber\\& =
                                                  \delta_{ik}\delta_{j\ell}
                                                  = \<a_i\otimes
                                                  b_j| a_k \otimes b_\ell\>_{\otimes},
	\labell{ecco}\; 
\end{align}
where we used the orthonormality of the bases and the fact that
$|a_i\otimes b_j\>$ is a basis of the tensor product space
${\cal A}\otimes{\cal B}$. Since the function $I$ is a linear
function that maps an orthonormal basis of ${\cal A}\otimes{\cal B}$
to an orthonormal basis of $\cal C$, $I$ is a an isomorphism (a
bijection that preserves the mathematical structure) between
${\cal A}\otimes{\cal B}$ and $\cal C$. As
${\cal C}\cong{\cal A}\otimes{\cal B}$ are isomorphic as Hilbert
spaces, they are mathematically equivalent: $c \in {\cal C}$ and $I^{-1}(c)$ (XXX) represent the same physical object. In this sense, we can loosely say that $I$ is the identity, as it connects spaces that are physically 
equivalent. So we can directly use the
tensor product to represent the composite state space. This means that
the map $m : \cal A \times \cal B \to \cal C$ is equivalent to the map
$\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$ in the
sense that $m \circ I^{-1} = \otimes$.$\square$

@@

\vskip 1\baselineskip A few comments on the proof: it is based on the
universal property of the tensor product, which uniquely characterizes
it. In step 1 we show that the bilinear map $m$ maps subsystems' bases
into the composite system basis. We also know that there exists a
tensor product map $ {T}=\otimes$ that can compose the vectors in
$\cal A$ and $\cal B$.  In step 2 we use the universal property: since
$m$ is a bilinear map, we are assured that there exists a unique
$I$ such that $I \circ T=m$. Since we show that $I$ is
an isomorphism, then $I$ bijectively maps vectors in $\cal C$
onto vectors in the tensor product space. Namely $m={T}=\otimes$. (XXX)
